---
title: "More String"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = TRUE, 
                      comment=NA, 
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 4, fig.width = 6, fig.align='center')

library(tidyverse) 
library(stringr)
library(babynames)
```


## Your turn 1

Make changes to the following code to ...

1. Isolate the last letter of every name

2. and create a logical variable that displays whether the last letter is one of "a", "e", "i", "o", "u", or "y".

3. Use a weighted mean to calculate the proportion of children whose name ends in a vowel, by year (see `?weighted.mean`).

4. and then display the results as a line plot.

```{r}
babynames %>%
  mutate(last = str_sub(____, ____),
         vowel = str_detect(____, _____)) %>%
  group_by(___) %>%
  summarize(p_vowel = weighted.mean(vowel, n)) %>%
  ggplot(aes(____, ____)) + 
  geom_line()
```

## Your turn 2

Determine how many baby names in 2017 contain "ders".

```{r}
babynames %>%
  filter(______) %>%
  pull(___) %>%
  ______(pattern = "ders") %>%
  sum()
```


## Your turn 3

### a. Fill in the code to determine how many baby names in 2017 began with a vowel.

```{r}
babynames %>% 
  filter(______) %>%                  # extract year 2017
  mutate(first.vowel = _______(name, ________)) %>%  # create logical column
  count(first.vowel)                    # create a frequency table
```


### b. Fill in the code to determine how many baby names in 2017 started or ended with a vowel.

```{r}
babynames %>% 
  filter(______) %>%           
  mutate(bookend.vowel = str_detect(name, "______________")) %>%  
  count(bookend.vowel)             
```

# Your Turn 4

## Consider the 2017 babynames data

### a. How many baby names contain "illie`?

```{r}
#number of illie
babynames %>% 
  filter(______) %>%
  summarize(
    num_illie = sum(_______(str_to_***(____), ""))
    )
# to get the actual names:
babynames %>% 
  filter(year == 2017, _______(str_to_***(), "")) 
```

### b. How many baby names start with "M"?

```{r}
# starts with M
babynames %>% 
  filter(______) %>%
  mutate(first_letter = _____(____, start = 1, end = 1)) %>%
  summarize(num_M = sum(first_letter == ""))
```


# Your turn 5

## The regular expression `"^[Ss](.*)(t+)(.+)(t+)"` detects "scuttlebutt", "Stetson", and "Scattter", but not "Scatter." Why?

We can verify the claim:

```{r}
x <- c("scuttlebutt", "Stetson", "Scattter", "Scatter")
str_detect(x, "_____________________")
```

Break it the regex down into chunks (if you don't see the answer right away). First we get the leading s or S:


```{r}
str_view(x, "___")
```

Then we get 0 or more values to follow:

```{r}
str_view(x, "^[Ss](.*)")
```
Now we get S/s through 1 or more t's in the string:

```{r}
str_view(x, "^[Ss](.*)(t+)")
```

Then we add 1 or more characters after the last t:
```{r}
str_view(x, "^[Ss](.*)(t+)(.+)")
```
But now we say that there must be 1 or more t's after the first "1 or more t, plus at least one other character" string. 

```{r}
str_view_all(x, "^[Ss](.*)(t+)(.+)(t+)")
str_detect(x, "^[Ss](.*)(t+)(.+)(t+)")
```

This is what rules `scatter` out because it has two t's but there isn't an extra character between the two. Stetson only has two t's but there are characters between the two t occurrences so it follows the rules. 

Finally, to get the entire word we need to end the regex with `(.*)`:

```{r}
str_view_all(x, "^[Ss](.*)(t+)(.+)(t+)___")
```

